\documentclass[11pt,a4paper]{report}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{tocloft}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{multicol}
\usepackage{float}

% Colors (matching software_docs.tex)
\definecolor{codebackground}{RGB}{245,245,245}
\definecolor{codecomment}{RGB}{0,128,0}
\definecolor{codekeyword}{RGB}{0,0,180}
\definecolor{codestring}{RGB}{163,21,21}
\definecolor{linkcolor}{RGB}{0,80,160}
\definecolor{chaptercolor}{RGB}{40,60,100}
\definecolor{pincolor}{RGB}{200,60,60}
\definecolor{muxcolor}{RGB}{60,120,200}
\definecolor{sensorcolor}{RGB}{60,160,60}
\definecolor{warnbg}{RGB}{255,240,230}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=linkcolor,
    urlcolor=linkcolor,
    citecolor=linkcolor,
    pdftitle={Kirin Hardware Integration Guide},
    pdfauthor={Strydr Silverberg},
}

% Code listing style
\lstdefinestyle{codestyle}{
    backgroundcolor=\color{codebackground},
    basicstyle=\ttfamily\small,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    commentstyle=\color{codecomment}\itshape,
    keywordstyle=\color{codekeyword}\bfseries,
    stringstyle=\color{codestring},
    numberstyle=\tiny\color{gray},
    numbers=left,
    numbersep=8pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=4,
    frame=single,
    framerule=0.5pt,
    rulecolor=\color{gray!50},
}
\lstset{style=codestyle}

% Chapter and section styling
\titleformat{\chapter}[display]
    {\normalfont\huge\bfseries\color{chaptercolor}}
    {\chaptertitlename\ \thechapter}{20pt}{\Huge}
\titlespacing*{\chapter}{0pt}{0pt}{30pt}

% Header and footer
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\leftmark}
\fancyhead[R]{\thepage}
\setlength{\headheight}{14pt}
\renewcommand{\headrulewidth}{0.4pt}

% Note box
\newcommand{\notebox}[1]{%
    \begin{center}
    \fcolorbox{gray}{gray!10}{%
        \parbox{0.9\textwidth}{\textbf{Note:} #1}
    }
    \end{center}
}

% Warning box
\newcommand{\warningbox}[1]{%
    \begin{center}
    \fcolorbox{red!70}{red!10}{%
        \parbox{0.9\textwidth}{\textbf{Warning:} #1}
    }
    \end{center}
}

% Tip box
\newcommand{\tipbox}[1]{%
    \begin{center}
    \fcolorbox{green!60!black}{green!5}{%
        \parbox{0.9\textwidth}{\textbf{Tip:} #1}
    }
    \end{center}
}

%------------------------------------------------------------------------------
% Document Information
%------------------------------------------------------------------------------
\title{%
    \vspace{-1cm}
    \rule{\textwidth}{1pt}\\[0.5cm]
    {\Huge\bfseries Kirin Autonomous Chess}\\[0.3cm]
    {\Large Hardware Integration Guide}\\[0.3cm]
    \rule{\textwidth}{1pt}\\[1cm]
    {\large Connecting the Raspberry Pi to the Physical Chess Board}
}
\author{Strydr Silverberg \& Colin Dake\\\\Colorado School of Mines\\2026 ProtoFund}
\date{\today}

\begin{document}

\maketitle
\thispagestyle{empty}

\vfill
\begin{center}
    \textit{A step-by-step guide to wiring, configuring, and testing the Kirin\\
    autonomous chess system on a Raspberry Pi.}
\end{center}
\vfill

\newpage
\tableofcontents
\newpage

%==============================================================================
\chapter{Overview}
%==============================================================================

\section{What This Document Covers}

This guide walks through the complete process of connecting the Kirin software to the physical chess board hardware. By the end, you will have a Raspberry Pi running the Kirin engine, reading 96 hall effect sensors through multiplexers, driving an OLED status display, reading operator buttons, and controlling a gantry system via GRBL over serial.

The system has four hardware subsystems that connect to the Pi:

\begin{enumerate}
    \item \textbf{Hall Effect Sensors (96 total)} --- detect piece presence on the board and in the two storage zones. Read through 6 multiplexers using 10 GPIO pins.
    \item \textbf{Operator UI} --- a 128$\times$64 I2C OLED for status messages plus Start, Stop, and Reset buttons on dedicated GPIO pins.
    \item \textbf{GRBL Gantry Controller} --- moves pieces on the physical board via an electromagnet on a 2-axis gantry. Communicates over USB serial.
    \item \textbf{Electromagnet} --- controlled by the GRBL board via the coolant output (M8/M9 commands).
\end{enumerate}

\section{System Architecture}

\begin{center}
\begin{tikzpicture}[
    box/.style={draw, rounded corners, minimum width=3cm, minimum height=1cm, align=center, font=\small},
    arr/.style={->, thick, >=stealth}
]
    % Pi
    \node[box, fill=blue!10] (pi) at (0, 0) {Raspberry Pi\\(Kirin Software)};
    
    % Sensors
    \node[box, fill=green!10] (mux) at (-5, -2.5) {6$\times$ CD74HC4067\\Multiplexers};
    \node[box, fill=green!10] (sensors) at (-5, -4.5) {96$\times$ A3144\\Hall Effect Sensors};
    \node[box, fill=blue!5] (ui) at (0, -3.5) {OLED Display\\+ Buttons};
    
    % GRBL
    \node[box, fill=orange!10] (grbl) at (5, -2.5) {GRBL Controller\\(Arduino/CNC Shield)};
    \node[box, fill=orange!10] (gantry) at (5, -4.5) {2-Axis Gantry\\+ Electromagnet};
    
    % Connections
    \draw[arr] (pi) -- node[left, font=\footnotesize] {10 GPIO pins} (mux);
    \draw[arr] (mux) -- node[left, font=\footnotesize] {16 channels each} (sensors);
    \draw[arr] (pi) -- node[right, font=\footnotesize] {I2C + 3 GPIO} (ui);
    \draw[arr] (pi) -- node[right, font=\footnotesize] {USB Serial} (grbl);
    \draw[arr] (grbl) -- node[right, font=\footnotesize] {Step/Dir + Spindle} (gantry);
\end{tikzpicture}
\end{center}

\section{Runtime Move Pipeline}

The physical game loop intentionally keeps a single software-to-hardware path for real hardware, dry-run, and simulation. For the common case where the human starts as white, one full cycle is:

\begin{enumerate}
    \item \textbf{Human moves a piece}: \texttt{BoardScanner::waitForLegalMove()} polls the board and storage muxes until the board or storage state differs from the synchronized baseline.
    \item \textbf{Sensors stabilize}: board and storage are debounced together. A candidate state must remain identical for \texttt{DEBOUNCE\_SCANS} reads separated by \texttt{DEBOUNCE\_DELAY\_MS}. The scanner never auto-commits after a timeout; unrecognized states only trigger an alert and the loop keeps waiting.
    \item \textbf{Move is matched to engine format}: \texttt{matchLegalMove()} compares the stable occupancy delta against the engine's legal move list. Captures are disambiguated with the exact storage slot that gained a piece and \texttt{PieceTracker}'s square-to-slot identity map.
    \item \textbf{Engine state advances}: the matched encoded move is passed to \texttt{makeMove()}, then \texttt{GameController::updateTracker()} keeps per-piece slot identity synchronized.
    \item \textbf{Engine responds}: \texttt{searchPosition()} selects \texttt{bestMove}. \texttt{GameController::executeEngineMove()} converts it to a \texttt{PhysicalMove}, builds a \texttt{MovePlan}, and sends that plan to the gantry controller.
    \item \textbf{Blockers move out of the way}: each \texttt{RelocationPlan} is executed with magnet on/off dwell time around the move.
    \item \textbf{Primary piece moves}: the engine-selected piece follows the planned path from source to destination. Captures require an exact storage slot before any capture G-code is emitted.
    \item \textbf{Blockers restore}: parked blockers are returned in reverse order. When a restoration path exists, the gantry follows it; only an empty restoration path falls back to a direct move.
    \item \textbf{Board is verified and re-baselined}: after successful hardware execution, the engine state, physical occupancy model, scanner board baseline, and scanner storage baseline are synchronized before waiting for the next human move.
\end{enumerate}

\notebox{The storage baseline used for a human turn is the last synchronized baseline, not a fresh scan taken after the prompt. This avoids a race where a fast player capture could accidentally become the new baseline instead of being detected as the move.}

\warningbox{Do not bypass \texttt{GameController::executeEngineMove()} for real games. It is the boundary that enforces exact capture-slot routing, special-move sequencing, and physical-board state updates. The lower-level G-code helpers are useful for tests and diagnostics only.}

\section{Prerequisites}

\begin{table}[H]
\centering
\begin{tabular}{ll}
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
Raspberry Pi & Pi 4 or Pi 5 (any RAM) \\
OS & Raspberry Pi OS (64-bit recommended) \\
Compiler & GCC with C++17 support \\
CMake & Version 3.10 or later \\
libgpiod & \texttt{sudo apt install libgpiod-dev} \\
Serial port & USB connection to GRBL controller \\
\bottomrule
\end{tabular}
\caption{Required components}
\end{table}


%==============================================================================
\chapter{Building the Software}
%==============================================================================

\section{Clone and Build}

\begin{lstlisting}[language=bash]
# Clone the repository
git clone https://github.com/strvdr/kirin-autonomous-chess.git
cd kirin-autonomous-chess/kirin

# Create build directory
mkdir build && cd build

# Configure with GPIO support
cmake .. -DCMAKE_BUILD_TYPE=Release \
         -DKIRIN_ENABLE_GPIOD=ON

# Build
cmake --build . --parallel
\end{lstlisting}

\warningbox{The \texttt{-DKIRIN\_ENABLE\_GPIOD=ON} option is essential on the Raspberry Pi. Without it, the board scanner and button controller compile in stub mode and GPIO calls are no-ops. The build will succeed either way, but sensors and buttons will not work without GPIO support enabled.}

\section{Install Dependencies}

\begin{lstlisting}[language=bash]
# Install libgpiod for GPIO access (no root required at runtime)
sudo apt install libgpiod-dev

# Verify installation
dpkg -l | grep libgpiod
\end{lstlisting}

\section{Run the Tests}

Before connecting any hardware, verify the software is correct:

\begin{lstlisting}[language=bash]
cd build

# Run all tests
ctest --output-on-failure

# Or run individually
./board_scanner_test     # Sensor move detection logic
./ui_controller_test     # OLED/button controller logic
./game_controller_test   # Engine-to-physical bridge
./board_interpreter_test # Path planning
./engine_test            # Engine validation
./captured_piece_test    # Capture encoding
\end{lstlisting}

All tests should pass. The board scanner tests run entirely in simulation mode and do not require GPIO hardware.


%==============================================================================
\chapter{Wiring the Hall Effect Sensors}
%==============================================================================

\section{Sensor Overview}

The system uses 96 A3144 hall effect sensors to detect the presence of magnets embedded in the chess pieces. The sensors are digital (output LOW when a magnet is detected, HIGH when no magnet is present) and require a 10K pull-up resistor each.

The 96 sensors are organized as follows:

\begin{table}[H]
\centering
\begin{tabular}{clcl}
\toprule
\textbf{Mux} & \textbf{Coverage} & \textbf{Channels} & \textbf{GPIO (SIG)} \\
\midrule
0 & Black storage zone, left of a-file & 0--15 & GPIO 5 \\
1 & Board files a--b & 0--15 & GPIO 6 \\
2 & Board files c--d & 0--15 & GPIO 13 \\
3 & Board files e--f & 0--15 & GPIO 19 \\
4 & Board files g--h & 0--15 & GPIO 26 \\
5 & White storage zone (16 slots) & 0--15 & GPIO 16 \\
\bottomrule
\end{tabular}
\caption{Multiplexer assignments}
\end{table}

\section{Multiplexer Wiring}

All six CD74HC4067 multiplexers share four address/select lines. The Pi sets these four lines (S0--S3) to select a channel (0--15), then reads the SIG output from each mux simultaneously.

\subsection{Shared Select Lines (Pi $\rightarrow$ All 6 Muxes)}

\begin{table}[H]
\centering
\begin{tabular}{ccc}
\toprule
\textbf{Select Line} & \textbf{Pi GPIO (BCM)} & \textbf{Mux Pin} \\
\midrule
S0 & GPIO 17 & Pin 10 (S0) on all muxes \\
S1 & GPIO 27 & Pin 11 (S1) on all muxes \\
S2 & GPIO 22 & Pin 14 (S2) on all muxes \\
S3 & GPIO 23 & Pin 15 (S3) on all muxes \\
\bottomrule
\end{tabular}
\caption{Select line wiring (active on all 6 muxes)}
\end{table}

\subsection{Signal Output Lines (Each Mux $\rightarrow$ Pi)}

Each mux has a single SIG output that goes to its own GPIO pin on the Pi:

\begin{table}[H]
\centering
\begin{tabular}{ccl}
\toprule
\textbf{Mux} & \textbf{SIG $\rightarrow$ Pi GPIO} & \textbf{Direction} \\
\midrule
0 & GPIO 5 & Input (Pi reads) \\
1 & GPIO 6 & Input \\
2 & GPIO 13 & Input \\
3 & GPIO 19 & Input \\
4 & GPIO 26 & Input \\
5 & GPIO 16 & Input \\
\bottomrule
\end{tabular}
\caption{Signal output wiring}
\end{table}

\subsection{Power and Enable}

\begin{itemize}
    \item Connect VCC on each mux to 3.3V (the Pi's GPIO operates at 3.3V).
    \item Connect GND on each mux to the Pi's GND.
    \item Connect the EN (enable) pin on each mux to GND (active-low enable, so grounding it keeps the mux always enabled).
\end{itemize}

\warningbox{Do not power the muxes from 5V. The CD74HC4067 will work at 3.3V, and connecting a 5V signal to the Pi's GPIO pins can damage the Pi.}

\section{Sensor-to-Channel Mapping}

Muxes 1--4 scan the 64 board squares. Each board mux covers two adjacent files. Channels 0--7 map to the first file in that pair from rank 8 down to rank 1; channels 8--15 map to the second file in that pair from rank 8 down to rank 1.

\begin{table}[H]
\centering
\begin{tabular}{ccl}
\toprule
\textbf{Mux} & \textbf{Files} & \textbf{Channel Pattern} \\
\midrule
1 & a--b & C0=a8, C1=a7, \ldots, C7=a1; C8=b8, \ldots, C15=b1 \\
2 & c--d & C0=c8, C1=c7, \ldots, C7=c1; C8=d8, \ldots, C15=d1 \\
3 & e--f & C0=e8, C1=e7, \ldots, C7=e1; C8=f8, \ldots, C15=f1 \\
4 & g--h & C0=g8, C1=g7, \ldots, C7=g1; C8=h8, \ldots, C15=h1 \\
\bottomrule
\end{tabular}
\caption{Board mux channel-to-square mapping}
\end{table}

Muxes 0 and 5 scan the storage zones. Mux 0 is black storage to the left of the a-file; mux 5 is white storage to the right of the h-file. Viewed from White's side, channels 0--7 are the left column top-to-bottom and channels 8--15 are the right column top-to-bottom. That means white storage shows back-rank labels in the left column and pawn labels in the right column, while black storage is visually flipped.

\begin{table}[H]
\centering
\begin{tabular}{cl|cl}
\toprule
\textbf{Channel} & \textbf{Inner Column Slot} & \textbf{Channel} & \textbf{Outer Column Slot} \\
\midrule
0 & R1 & 8 & P1 \\
1 & R2 & 9 & P2 \\
2 & B1 & 10 & P3 \\
3 & B2 & 11 & P4 \\
4 & N1 & 12 & P5 \\
5 & N2 & 13 & P6 \\
6 & Q & 14 & P7 \\
7 & K & 15 & P8 \\
\bottomrule
\end{tabular}
\caption{Storage mux channel-to-slot mapping for both mux 0 and mux 5}
\end{table}

\subsection{Full Board Sensor Map}

The diagram below shows the complete mapping from every board square to its mux and channel. Each cell shows \textbf{M\textit{x}:C\textit{y}} where \textit{x} is the mux number (1--4) and \textit{y} is the channel (0--15). The mux select address (S3 S2 S1 S0) is shown for each channel on the right.

\begin{center}
\begin{tikzpicture}[scale=0.85]
    % Colors for each mux
    \definecolor{mux1}{RGB}{200,220,255}
    \definecolor{mux2}{RGB}{200,255,220}
    \definecolor{mux3}{RGB}{255,240,200}
    \definecolor{mux4}{RGB}{255,210,210}
    
    % Draw the 8x8 grid
    \foreach \col in {0,...,7} {
        \foreach \row in {0,...,7} {
            % Determine mux number from file pair.
            \pgfmathtruncatemacro{\muxnum}{int(\col/2) + 1}
            % Determine channel: first file C0-C7 rank 8-1, second file C8-C15 rank 8-1.
            \pgfmathtruncatemacro{\filemod}{mod(\col,2)}
            \pgfmathtruncatemacro{\chan}{\filemod*8 + \row}
            
            % Pick color based on mux
            \ifnum\muxnum=1 \def\cellcolor{mux1}\fi
            \ifnum\muxnum=2 \def\cellcolor{mux2}\fi
            \ifnum\muxnum=3 \def\cellcolor{mux3}\fi
            \ifnum\muxnum=4 \def\cellcolor{mux4}\fi
            
            % Draw cell
            \fill[\cellcolor] (\col*2, -\row*1.2) rectangle ++(2, -1.2);
            \draw[gray] (\col*2, -\row*1.2) rectangle ++(2, -1.2);
            
            % Mux:Channel label
            \node[font=\footnotesize\bfseries] at (\col*2+1, -\row*1.2-0.4) {M\muxnum:C\chan};
            
            % Square name
            \pgfmathtruncatemacro{\rank}{8 - \row}
            \node[font=\scriptsize, gray] at (\col*2+1, -\row*1.2-0.85) {%
                \ifcase\col a\or b\or c\or d\or e\or f\or g\or h\fi\rank};
        }
    }
    
    % File labels (top)
    \foreach \col/\file in {0/a, 1/b, 2/c, 3/d, 4/e, 5/f, 6/g, 7/h} {
        \node[font=\small\bfseries] at (\col*2+1, 0.4) {\file};
    }
    
    % Rank labels (left)
    \foreach \row in {0,...,7} {
        \pgfmathtruncatemacro{\rank}{8 - \row}
        \node[font=\small\bfseries] at (-0.5, -\row*1.2-0.6) {\rank};
    }
    
    % Mux labels (right side)
    \node[font=\small\bfseries, fill=mux1, draw=gray, rounded corners, minimum width=2.2cm] 
        at (18.5, -0.6-0.6) {Mux 1 a--b};
    \node[font=\small\bfseries, fill=mux2, draw=gray, rounded corners, minimum width=2.2cm] 
        at (18.5, -2.4-0.6) {Mux 2 c--d};
    \node[font=\small\bfseries, fill=mux3, draw=gray, rounded corners, minimum width=2.2cm] 
        at (18.5, -4.8-0.6) {Mux 3 e--f};
    \node[font=\small\bfseries, fill=mux4, draw=gray, rounded corners, minimum width=2.2cm] 
        at (18.5, -7.2-0.6) {Mux 4 g--h};
    
    % Select address reference (bottom)
    \node[anchor=north west, font=\small, text width=16cm] at (0, -10.5) {%
        \textbf{Select address (S3 S2 S1 S0) for each channel:}\\[2pt]
        \begin{tabular}{cccccccc}
        C0: 0000 & C1: 0001 & C2: 0010 & C3: 0011 & C4: 0100 & C5: 0101 & C6: 0110 & C7: 0111 \\
        C8: 1000 & C9: 1001 & C10: 1010 & C11: 1011 & C12: 1100 & C13: 1101 & C14: 1110 & C15: 1111
        \end{tabular}
    };
\end{tikzpicture}
\end{center}

\subsection{Storage Zone Sensor Map}

The two storage zone muxes each have 16 channels mapped to exact piece storage slots. The black storage zone is left of the a-file on mux 0. The white storage zone is right of the h-file on mux 5. Channels 0--7 are always the visual left column top-to-bottom and channels 8--15 are the visual right column top-to-bottom. White storage therefore maps C0--C7 to R1, R2, B1, B2, N1, N2, Q, K and C8--C15 to P1--P8, while black storage is flipped with C0--C7 = P1--P8 and C8--C15 = R1, R2, B1, B2, N1, N2, Q, K.

\begin{center}
\begin{tikzpicture}[scale=0.85]
    \definecolor{storebk}{RGB}{255,210,210}
    \definecolor{storewh}{RGB}{200,220,255}
    
    % Black storage (left)
    \node[font=\large\bfseries] at (3.5, 1) {Black Storage --- Mux 0 (GPIO 5)};
    \foreach \slot/\name in {0/P1,1/P2,2/P3,3/P4,4/P5,5/P6,6/P7,7/P8} {
        \fill[storebk] (0, -\slot*0.9) rectangle ++(3.5, -0.9);
        \draw[gray] (0, -\slot*0.9) rectangle ++(3.5, -0.9);
        \node[font=\footnotesize] at (1.75, -\slot*0.9-0.45) {C\slot: \name};
    }
    \foreach \slot/\name in {0/R1,1/R2,2/B1,3/B2,4/N1,5/N2,6/Q,7/K} {
        \pgfmathtruncatemacro{\chan}{\slot+8}
        \fill[storebk!60] (4, -\slot*0.9) rectangle ++(3.5, -0.9);
        \draw[gray] (4, -\slot*0.9) rectangle ++(3.5, -0.9);
        \node[font=\footnotesize] at (5.75, -\slot*0.9-0.45) {C\chan: \name};
    }
    
    % White storage (right)
    \node[font=\large\bfseries] at (14.5, 1) {White Storage --- Mux 5 (GPIO 16)};
    \foreach \slot/\name in {0/R1,1/R2,2/B1,3/B2,4/N1,5/N2,6/Q,7/K} {
        \fill[storewh] (11, -\slot*0.9) rectangle ++(3.5, -0.9);
        \draw[gray] (11, -\slot*0.9) rectangle ++(3.5, -0.9);
        \node[font=\footnotesize] at (12.75, -\slot*0.9-0.45) {C\slot: \name};
    }
    \foreach \slot/\name in {0/P1,1/P2,2/P3,3/P4,4/P5,5/P6,6/P7,7/P8} {
        \pgfmathtruncatemacro{\chan}{\slot+8}
        \fill[storewh!60] (15, -\slot*0.9) rectangle ++(3.5, -0.9);
        \draw[gray] (15, -\slot*0.9) rectangle ++(3.5, -0.9);
        \node[font=\footnotesize] at (16.75, -\slot*0.9-0.45) {C\chan: \name};
    }
    
    \node[font=\footnotesize, text width=18cm, anchor=north west] at (0, -7.8) {%
        \textbf{Black storage left/right columns:} C0--C7 = P1, P2, P3, P4, P5, P6, P7, P8; C8--C15 = R1, R2, B1, B2, N1, N2, Q, K.\\
        \textbf{White storage left/right columns:} C0--C7 = R1, R2, B1, B2, N1, N2, Q, K; C8--C15 = P1, P2, P3, P4, P5, P6, P7, P8.
    };
\end{tikzpicture}
\end{center}

\tipbox{If your physical wiring doesn't match this channel order, the software's idea of the board will diverge from the hardware. Either re-wire the sensors to this map or update \texttt{toSquareIndex()} and the storage-slot mapping in \texttt{board\_scanner.cpp}.}

\section{A3144 Sensor Circuit}

Each A3144 sensor has three pins: VCC, GND, and OUTPUT. The output is open-collector, so it needs a pull-up resistor:

\begin{center}
\begin{tikzpicture}[scale=0.9]
    % VCC rail
    \draw[thick] (0, 4) -- (6, 4) node[right] {3.3V};
    
    % Sensor
    \draw[thick] (1, 4) -- (1, 2.5);
    \node[draw, minimum width=1.5cm, minimum height=1cm, fill=sensorcolor!20] at (1, 2) {A3144};
    \draw[thick] (1, 1.5) -- (1, 0);
    \draw[thick] (1.75, 2) -- (3, 2);
    
    % Pull-up resistor
    \draw[thick] (3, 4) -- (3, 3.2);
    \node[draw, minimum width=0.4cm, minimum height=0.8cm, fill=white] at (3, 2.8) {\footnotesize 10K};
    \draw[thick] (3, 2.4) -- (3, 2);
    
    % To mux
    \draw[thick, ->] (3, 2) -- (5.5, 2) node[right] {To mux channel};
    
    % GND rail
    \draw[thick] (0, 0) -- (6, 0) node[right] {GND};
    
    % Labels
    \node[above] at (1, 4) {\footnotesize VCC};
    \node[below] at (1, 0) {\footnotesize GND};
    \node[above] at (3, 4) {\footnotesize Pull-up};
    \node[below] at (3, 2) {\footnotesize OUT};
\end{tikzpicture}
\end{center}

When a magnet is present, the A3144 pulls its output LOW. When no magnet is present, the pull-up resistor holds the output HIGH. The software inverts this logic: LOW = piece present, HIGH = empty.


%==============================================================================
\chapter{Connecting the Gantry}
%==============================================================================

\section{GRBL Connection}

The gantry controller (typically an Arduino Uno with a CNC shield running GRBL firmware) connects to the Pi via USB. This appears as a serial port, usually \texttt{/dev/ttyUSB0} or \texttt{/dev/ttyACM0}.

\begin{lstlisting}[language=bash]
# Find the serial port
ls /dev/ttyUSB* /dev/ttyACM*

# Check permissions (your user must be in the 'dialout' group)
groups $USER | grep dialout

# If not in the group:
sudo usermod -aG dialout $USER
# Log out and back in for this to take effect
\end{lstlisting}

\section{GRBL Settings}

The Kirin software communicates at 115200 baud (GRBL default). Verify the GRBL settings match the physical gantry dimensions. Connect to GRBL with a serial terminal to check:

\begin{lstlisting}[language=bash]
# Connect with minicom (or screen, picocom, etc.)
minicom -D /dev/ttyUSB0 -b 115200

# In the GRBL console, type:
$$
# This prints all settings
\end{lstlisting}

Key settings to verify:

\begin{table}[H]
\centering
\begin{tabular}{clp{7cm}}
\toprule
\textbf{Setting} & \textbf{Parameter} & \textbf{Notes} \\
\midrule
\$100 & X steps/inch & Must match your stepper + lead screw \\
\$101 & Y steps/inch & Must match your stepper + lead screw \\
\$110 & X max rate (in/min) & Software uses 1000 in/min default \\
\$111 & Y max rate (in/min) & Must be $\geq$ 1000 \\
\$120 & X acceleration (in/s$^2$) & Start with 10, increase after testing \\
\$121 & Y acceleration (in/s$^2$) & Start with 10, increase after testing \\
\$130 & X max travel (in) & Should cover the calibrated command envelope (currently up to X=19.0) \\
\$131 & Y max travel (in) & Should cover the calibrated command envelope (currently up to Y=14.92) \\
\$22 & Homing cycle & Set to 1 (enabled) \\
\bottomrule
\end{tabular}
\caption{Critical GRBL settings}
\end{table}

\section{Electromagnet Control}

The electromagnet is controlled through GRBL's coolant output:

\begin{itemize}
    \item \texttt{M8} = Magnet ON (coolant enable --- repurposed for electromagnet)
    \item \texttt{M9} = Magnet OFF (coolant disable)
\end{itemize}

The CNC shield's coolant enable pin drives a MOSFET or relay that controls the electromagnet. Verify this works before running any game:

\begin{lstlisting}[language=bash]
# In the GRBL console:
M8     # Magnet should engage
G4 P2.000  # Wait 2 seconds; GRBL P values are seconds
M9     # Magnet should release
\end{lstlisting}

\notebox{Kirin's \texttt{dwell(milliseconds)} helper accepts milliseconds but emits GRBL seconds. For example, the default pick/release settle time \texttt{dwell(100)} is sent as \texttt{G4 P0.100}. Sending \texttt{G4 P100} would pause for 100 seconds on GRBL, not 100 milliseconds.}

\section{GRBL Motion Mode Setup}

After homing, Kirin sends a deterministic modal setup sequence before normal movement:

\begin{lstlisting}[language=bash]
$H     # Home gantry
G20    # Use inches
G90    # Use absolute coordinates
G94    # Feed rate is units per minute
M9     # Ensure magnet is off
\end{lstlisting}

\warningbox{Do not assume the GRBL controller is already in the correct modal state from a previous session. The software's coordinate constants and feed rates are expressed in inches and absolute G54 work coordinates, so \texttt{G20}, \texttt{G90}, and \texttt{G94} must be active before moving pieces.}

\section{Physical Coordinate System}

The software maps board squares to physical coordinates in inches. The board layout is:

\begin{center}
\begin{tikzpicture}[scale=0.48]
    % Rail
    \draw[thick] (0, 0) rectangle (20.5, 19);
    \draw[<->] (0.5, 17.5) -- (20, 17.5) node[midway, above] {\footnotesize 22'' physical rail travel};
    
    % Board
    \draw[thick, fill=blue!5] (4.25, 3.67) rectangle (16.25, 15.67);
    \node at (10.25, 10.0) {\large 12'' $\times$ 12'' Board};
    \node at (10.25, 8.7) {\footnotesize 8$\times$8 grid, 1.5'' squares};
    
    % Black storage
    \draw[thick, fill=red!5] (0.75, 3.67) rectangle (3.75, 15.67);
    \node[rotate=90] at (2.25, 10.0) {\footnotesize Black Storage};
    
    % White storage
    \draw[thick, fill=green!5] (16.75, 3.67) rectangle (19.75, 15.67);
    \node[rotate=90] at (18.25, 10.0) {\footnotesize White Storage};
    
    % Dimensions
    \draw[<->] (0.75, 1.3) -- (19.75, 1.3) node[midway, below] {\footnotesize 19'' occupied hardware width};
    \draw[<->] (4.25, 2.6) -- (16.25, 2.6) node[midway, below] {\footnotesize 12'' board};
    \draw[<->] (0.75, 3.0) -- (3.75, 3.0) node[midway, below] {\footnotesize storage};
    \draw[<->] (16.75, 3.0) -- (19.75, 3.0) node[midway, below] {\footnotesize storage};
    \draw[<->] (21.5, 3.67) -- (21.5, 15.67) node[midway, right] {\footnotesize 12''};
    
    % a1 marker
    \filldraw[red] (5.0, 4.42) circle (0.12);
    \node[red, above right] at (5.0, 4.42) {\footnotesize a1 (5.000, 4.420)};
    
    % h8 marker
    \filldraw[red] (15.5, 14.92) circle (0.12);
    \node[red, below left] at (15.5, 14.92) {\footnotesize h8 (15.500, 14.920)};
\end{tikzpicture}
\end{center}

The physical rails are 22'' long. The occupied hardware width is 19'': 3'' storage + 0.5'' gap + 12'' board + 0.5'' gap + 3'' storage. The coordinate system is the calibrated G54 work frame used by Kirin's normal \texttt{G1} moves. Square a1 is centered at (5.000'', 4.420'') and each square is 1.5'' wide. Ranks advance along +X and files advance along +Y, so h8 is centered at (15.500'', 14.920''). The storage rows use the same 1.5'' vertical spacing and align with the board file coordinates.


%==============================================================================
\chapter{Operator UI: OLED Display and Buttons}
%==============================================================================

\section{OLED Display}

Kirin supports a 128$\times$64 SSD1306-compatible OLED on the Raspberry Pi's I2C bus. The default software configuration is \texttt{/dev/i2c-1} at address \texttt{0x3C}.

\begin{table}[H]
\centering
\begin{tabular}{cll}
\toprule
\textbf{Signal} & \textbf{BCM GPIO / Header Pin} & \textbf{Connects To} \\
\midrule
SDA & GPIO 2 / Pin 3 & OLED SDA \\
SCL & GPIO 3 / Pin 5 & OLED SCL \\
3.3V & Pin 1 or Pin 17 & OLED VCC \\
GND & Any Pi ground pin & OLED GND \\
\bottomrule
\end{tabular}
\caption{OLED I2C wiring}
\end{table}

The display controller is implemented in \texttt{display\_controller.h/.cpp}. Physical mode uses it for high-level status messages: booting, hardware ready, waiting for the human move, engine thinking, executing a move, illegal board state, wrong storage slot, game over, and idle. Use \texttt{--oled-test} to initialize the display and show a test screen without starting a physical game.

\section{Operator Buttons}

Kirin supports three active-low buttons using Raspberry Pi internal pull-ups. Pressing a button connects its GPIO line to ground. The button controller is implemented in \texttt{button\_controller.h/.cpp}.

\begin{table}[H]
\centering
\begin{tabular}{cll}
\toprule
\textbf{Button} & \textbf{BCM GPIO / Header Pin} & \textbf{Meaning at Safe Prompt} \\
\midrule
Start & GPIO 20 / Pin 38 & Idle: start \texttt{play white}; active game: run \texttt{go} \\
Stop & GPIO 21 / Pin 40 & Active game: stop game; idle: exit physical mode \\
Reset & GPIO 24 / Pin 18 & Home the gantry \\
\bottomrule
\end{tabular}
\caption{Operator button wiring and behavior}
\end{table}

\warningbox{The buttons are high-level UI controls, not an asynchronous emergency stop. They are polled only at safe physical-mode prompts and are not polled during blocking gantry motion, engine search, or \texttt{waitForHumanMove}.}


%==============================================================================
\chapter{First Power-On Procedure}
%==============================================================================

\section{Step 1: Verify GPIO Chip}

The Pi 5 and Pi 4 use different GPIO chip paths. The software auto-detects, but verify:

\begin{lstlisting}[language=bash]
# Pi 5:
ls /dev/gpiochip4  # Should exist

# Pi 4 and earlier:
ls /dev/gpiochip0  # Should exist

# List all GPIO chips:
gpiodetect
\end{lstlisting}

\section{Step 2: Start Kirin in Physical Mode}

\begin{lstlisting}[language=bash]
cd build
./kirin --physical /dev/ttyUSB0
\end{lstlisting}

You should see:

\begin{lstlisting}[language=bash, numbers=none]
Kirin Autonomous Chess System v0.3
Physical Board Mode

[DISPLAY] OLED initialized on /dev/i2c-1 address 0x3C
[BUTTONS] Start=GPIO20 Stop=GPIO21 Reset=GPIO24
Connecting to gantry on /dev/ttyUSB0...
Connected to gantry on /dev/ttyUSB0
Initializing board scanner...
[SCANNER] GPIO initialized successfully (6 muxes: 4 board + 2 storage)
  Select lines: GPIO 17, 27, 22, 23
  Board muxes:  GPIO 6, 13, 19, 26
  Storage muxes: GPIO 5 (black), 16 (white)
Board scanner initialized
Homing gantry...
Homing complete

Ready. Type 'help' for commands.
\end{lstlisting}

If the scanner fails to initialize, check your wiring and ensure \texttt{-DHAS\_GPIOD} was used during compilation.

\section{Step 3: Run Sensor Diagnostics}

\begin{lstlisting}[language=bash, numbers=none]
kirin> diag
\end{lstlisting}

This reads every sensor individually and prints the result. Walk through each square with a magnet (or chess piece) and verify:

\begin{itemize}
    \item Each channel maps to the correct square
    \item ``PIECE'' appears when the magnet is present, ``empty'' when removed
    \item No stuck sensors (always reading PIECE or always empty)
    \item Storage zone sensors respond correctly
\end{itemize}

\section{Step 4: Quick Scan}

Place all 32 pieces on the board in the starting position, then:

\begin{lstlisting}[language=bash, numbers=none]
kirin> scan
\end{lstlisting}

You should see:

\begin{lstlisting}[language=bash, numbers=none]
  Sensor Scan:
    a b c d e f g h
  8 # # # # # # # # 8
  7 # # # # # # # # 7
  6 . . . . . . . . 6
  5 . . . . . . . . 5
  4 . . . . . . . . 4
  3 . . . . . . . . 3
  2 # # # # # # # # 2
  1 # # # # # # # # 1
    a b c d e f g h
  Pieces detected: 32
\end{lstlisting}

If any squares are wrong, re-check the wiring for that mux/channel.

\section{Step 5: Test the Gantry}

\begin{lstlisting}[language=bash, numbers=none]
kirin> test
\end{lstlisting}

This moves the gantry to e4, then d5, engages and disengages the magnet, and returns home. Watch for:

\begin{itemize}
    \item Smooth motion with no skipped steps
    \item Correct positioning over the square centers
    \item Magnet picks up and releases cleanly
    \item No collisions with pieces or board edges
\end{itemize}

\section{Step 6: Gantry Coordinate Calibration}

If the gantry doesn't align with the square centers, you need to adjust either the GRBL steps/inch settings or the software constants. Place a piece on a1 and command:

\begin{lstlisting}[language=bash, numbers=none]
kirin> home
\end{lstlisting}

Then type \texttt{test}. The gantry should move to e4 (center of the board). If it's offset, adjust \texttt{A1\_CENTER\_X} and \texttt{A1\_CENTER\_Y} in \texttt{gantry\_controller.h}, or recalibrate your GRBL \$100/\$101 steps/inch settings.

\tipbox{Get a1 and h8 aligned first. Since the coordinate system is linear, if these two corners are correct, all 64 squares will be correct.}


%==============================================================================
\chapter{Running Your First Game}
%==============================================================================

\section{Stage 1: Manual Move Entry (No Sensors)}

Start with manual move input to verify the gantry executes moves correctly:

\begin{lstlisting}[language=bash, numbers=none]
kirin> newgame black
\end{lstlisting}

This sets up the board (gantry moves all pieces from storage to starting squares), then the engine plays black. You type your moves:

\begin{lstlisting}[language=bash, numbers=none]
kirin> move e2e4
kirin> move d2d4
kirin> move g1f3
\end{lstlisting}

After each of your moves, the engine will think and physically move a piece in response. Watch for correct piece movement, especially during captures (the captured piece should go to the correct storage slot).

\section{Stage 2: Sensor-Based Play}

Once manual mode works, switch to the full sensor experience:

\begin{lstlisting}[language=bash, numbers=none]
kirin> play black
\end{lstlisting}

Now make your moves by physically moving pieces on the board. The system will detect your move automatically. The terminal will show:

\begin{lstlisting}[language=bash, numbers=none]
Move 1. -- Your turn. Make your move on the board.
Waiting for your move...
Detected move: e2e4
\end{lstlisting}

\subsection{What to Test}

\begin{enumerate}
    \item \textbf{Normal moves}: Pick up a pawn, place it on a new square.
    \item \textbf{Pondering}: Pick up a piece, hold it for 10+ seconds, put it back. The system should \emph{not} commit a move.
    \item \textbf{Take-back}: Pick up one piece, put it back, pick up a different piece and move it. Only the final move should register.
    \item \textbf{Captures}: Pick up the opponent's piece, place it in the storage zone, then move your piece to the vacated square.
    \item \textbf{Castling}: Move the king two squares, then move the rook. The system waits until both pieces are in their final positions.
    \item \textbf{Illegal state}: Move a piece to an illegal square. After 10 seconds, the system will alert you.
\end{enumerate}

\section{Stage 3: Full Game}

Play a complete game to checkmate or stalemate. The system will announce the result:

\begin{lstlisting}[language=bash, numbers=none]
==============================
  CHECKMATE -- White wins!
==============================
\end{lstlisting}


%==============================================================================
\chapter{Troubleshooting}
%==============================================================================

\section{Sensor Issues}

\begin{table}[H]
\centering
\begin{tabular}{p{5cm}p{9cm}}
\toprule
\textbf{Symptom} & \textbf{Solution} \\
\midrule
All sensors read empty & Check mux power (3.3V), verify EN pin is grounded, check select line wiring \\
\addlinespace
One mux reads all empty & Check that mux's SIG output wire to the Pi GPIO pin \\
\addlinespace
Random noise / flickering & Increase \texttt{DEBOUNCE\_SCANS} (default 3) or \texttt{DEBOUNCE\_DELAY\_MS} (default 50ms) in \texttt{board\_scanner.h} \\
\addlinespace
Sensors read wrong squares & Your channel-to-square wiring doesn't match the expected layout. Run \texttt{diag} and compare against the mapping table. Either re-wire or modify \texttt{toSquareIndex()} \\
\addlinespace
Sensors trigger from gantry magnet & The gantry electromagnet is interfering with nearby hall sensors. Add shielding around the gantry magnet, increase debounce, or only scan when the gantry is idle \\
\addlinespace
Scanner init fails & Verify \texttt{libgpiod-dev} is installed and the build used \texttt{-DHAS\_GPIOD}. Check that the GPIO chip exists (\texttt{/dev/gpiochip4} on Pi 5, \texttt{/dev/gpiochip0} on Pi 4) \\
\bottomrule
\end{tabular}
\caption{Common sensor issues}
\end{table}

\section{Gantry Issues}

\begin{table}[H]
\centering
\begin{tabular}{p{5cm}p{9cm}}
\toprule
\textbf{Symptom} & \textbf{Solution} \\
\midrule
``Failed to connect to gantry'' & Check USB cable, verify serial port path (\texttt{ls /dev/ttyUSB*}), ensure user is in \texttt{dialout} group \\
\addlinespace
Homing fails & Verify limit switches are connected and GRBL setting \$22=1 (homing enabled) \\
\addlinespace
Gantry moves to wrong position & Recalibrate GRBL steps/inch (\$100, \$101) or adjust \texttt{A1\_CENTER\_X}/\texttt{A1\_CENTER\_Y} in the software \\
\addlinespace
Magnet doesn't engage & Check GRBL coolant output wiring, test with \texttt{M8}/\texttt{M9} in a serial console \\
\addlinespace
Pieces dragged/knocked over & Reduce \texttt{FEED\_RATE} (default 1000 in/min), verify magnet strength is appropriate \\
\addlinespace
GRBL alarm triggered & The software auto-resets on alarm. Check error description in terminal output. Common cause: soft limits exceeded (piece near board edge) \\
\bottomrule
\end{tabular}
\caption{Common gantry issues}
\end{table}

\section{Move Detection Issues}

\begin{table}[H]
\centering
\begin{tabular}{p{5cm}p{9cm}}
\toprule
\textbf{Symptom} & \textbf{Solution} \\
\midrule
System doesn't detect my move & Verify the sensor under the destination square works (\texttt{diag}). Check that you placed the captured piece in storage (required for capture detection). \\
\addlinespace
System hangs waiting for move & If it's a rare ambiguous capture (queen can capture two different pieces from the same square), fall back to typed input: press Ctrl+C and use \texttt{move} command. \\
\addlinespace
``Board state not recognized'' alert & You've left the board in a state that doesn't correspond to any legal move for 10+ seconds. Fix piece placement. \\
\addlinespace
Board mismatch after engine move & The gantry dropped a piece. The system will show which squares are wrong. Fix the pieces manually and type \texttt{continue}. \\
\bottomrule
\end{tabular}
\caption{Common move detection issues}
\end{table}


%==============================================================================
\chapter{GPIO Pin Reference}
%==============================================================================

Complete pin assignment table using BCM (Broadcom) numbering:

\begin{table}[H]
\centering
\begin{tabular}{ccll}
\toprule
\textbf{BCM Pin} & \textbf{Direction} & \textbf{Function} & \textbf{Connects To} \\
\midrule
17 & Output & Select S0 & S0 on all 6 muxes \\
27 & Output & Select S1 & S1 on all 6 muxes \\
22 & Output & Select S2 & S2 on all 6 muxes \\
23 & Output & Select S3 & S3 on all 6 muxes \\
\addlinespace
5  & Input  & Mux 0 SIG & Black storage zone \\
6  & Input  & Mux 1 SIG & Board files a--b \\
13 & Input  & Mux 2 SIG & Board files c--d \\
19 & Input  & Mux 3 SIG & Board files e--f \\
26 & Input  & Mux 4 SIG & Board files g--h \\
16 & Input  & Mux 5 SIG & White storage zone \\
\addlinespace
2  & I2C & OLED SDA & OLED display data \\
3  & I2C & OLED SCL & OLED display clock \\
\addlinespace
20 & Input & Start button & Active-low, pull-up enabled \\
21 & Input & Stop button & Active-low, pull-up enabled \\
24 & Input & Reset button & Active-low, pull-up enabled \\
\bottomrule
\end{tabular}
\caption{All GPIO pin assignments for sensors and operator UI}
\end{table}

\notebox{The mux pin assignments can be overridden by passing custom arrays to the \texttt{BoardScanner} constructor. The button defaults are defined in \texttt{button\_controller.h}; the OLED defaults are defined in \texttt{display\_controller.h}. If you need to change them, update the corresponding default constants and rebuild.}

\begin{table}[H]
\centering
\begin{tabular}{cl}
\toprule
\textbf{Raspberry Pi Header Pin} & \textbf{BCM GPIO} \\
\midrule
Pin 11 & GPIO 17 (S0) \\
Pin 13 & GPIO 27 (S1) \\
Pin 15 & GPIO 22 (S2) \\
Pin 16 & GPIO 23 (S3) \\
Pin 3 & GPIO 2 (OLED SDA) \\
Pin 5 & GPIO 3 (OLED SCL) \\
Pin 29 & GPIO 5 (Mux 0) \\
Pin 31 & GPIO 6 (Mux 1) \\
Pin 33 & GPIO 13 (Mux 2) \\
Pin 35 & GPIO 19 (Mux 3) \\
Pin 37 & GPIO 26 (Mux 4) \\
Pin 36 & GPIO 16 (Mux 5) \\
Pin 38 & GPIO 20 (Start button) \\
Pin 40 & GPIO 21 (Stop button) \\
Pin 18 & GPIO 24 (Reset button) \\
\bottomrule
\end{tabular}
\caption{BCM to physical header pin mapping}
\end{table}


%==============================================================================
\chapter{Software Command Reference}
%==============================================================================

\section{Command-Line Validation Modes}

Use these modes before powering the gantry:

\begin{table}[H]
\centering
\begin{tabular}{p{5cm}p{9cm}}
\toprule
\textbf{Mode} & \textbf{Purpose} \\
\midrule
\texttt{--simulate} & Interactive move-by-move mode. It uses the same \texttt{GameController} to \texttt{GrblController} dry-run execution path as physical mode, but prints G-code instead of sending it to serial hardware. \\
\addlinespace
\texttt{--dryrun [MOVES [DEPTH]]} & Engine-vs-engine dry run. It exercises setup, path planning, capture-slot routing, blocker relocation, and restoration G-code without connected hardware. \\
\bottomrule
\end{tabular}
\caption{Command-line validation modes}
\end{table}

\section{Physical Mode Commands}

Quick reference for all commands available in physical mode:

\begin{table}[H]
\centering
\begin{tabular}{p{4.5cm}p{9.5cm}}
\toprule
\textbf{Command} & \textbf{Description} \\
\midrule
\texttt{play [white|black]} & Start a sensor-based game. Engine plays the specified color (default: white). Moves are detected automatically via hall effect sensors. \\
\addlinespace
\texttt{newgame [white|black]} & Start a game with manual (typed) move entry. Useful as a fallback if sensors aren't working. \\
\addlinespace
\texttt{move <uci>} & Manually enter a move in UCI format (e.g., \texttt{move e2e4}, \texttt{move e7e8q} for promotion). \\
\addlinespace
\texttt{go} & Engine makes one move. \\
\addlinespace
\texttt{scan} & Scan and display all 96 sensors (board + storage zones). \\
\addlinespace
\texttt{diag} & Run full sensor diagnostic --- reads each sensor individually with channel-by-channel output. \\
\addlinespace
\texttt{board} & Display the engine's internal board position. \\
\addlinespace
\texttt{physical} & Display the physical board occupancy tracking. \\
\addlinespace
\texttt{fen <fen>} & Set position from a FEN string. \\
\addlinespace
\texttt{home} & Home the gantry to (0,0). \\
\addlinespace
\texttt{test} & Run gantry hardware test (move to e4, d5, test magnet). \\
\addlinespace
\texttt{quit} & Exit the program. \\
\bottomrule
\end{tabular}
\caption{Physical mode commands}
\end{table}

\notebox{The \texttt{--oled-test} command-line flag starts only the OLED test screen. It is useful for verifying display wiring before testing the full physical board. The Start, Stop, and Reset buttons mirror safe prompt actions as documented in the Operator UI chapter.}

%==============================================================================
\chapter{Tunable Parameters}
%==============================================================================

These constants in \texttt{board\_scanner.h} and \texttt{gantry\_controller.h} may need adjustment based on your specific hardware:

\begin{table}[H]
\centering
\begin{tabular}{p{4.5cm}cp{7cm}}
\toprule
\textbf{Constant} & \textbf{Default} & \textbf{Purpose} \\
\midrule
\texttt{MUX\_SETTLE\_US} & 10 $\mu$s & Time to wait after changing mux address before reading. Increase if readings are noisy. \\
\addlinespace
\texttt{DEBOUNCE\_SCANS} & 3 & Number of consecutive identical scans required before accepting a reading as stable. \\
\addlinespace
\texttt{DEBOUNCE\_DELAY\_MS} & 50 ms & Delay between debounce scans. Total debounce time = \texttt{SCANS} $\times$ \texttt{DELAY}. \\
\addlinespace
\texttt{POLL\_INTERVAL\_MS} & 20 ms & How often to poll sensors during \texttt{waitForLegalMove}. \\
\addlinespace
\texttt{ILLEGAL\_STATE\_ALERT\_MS} & 10,000 ms & How long the board can sit in an unrecognized state before alerting the player. \\
\addlinespace
\texttt{FEED\_RATE} & 1000 in/min & Gantry movement speed. Reduce if the gantry skips steps or knocks pieces over. \\
\addlinespace
\texttt{dwell(100)} & 100 ms & Pick/release settle time around magnet on/off. Encoded as \texttt{G4 P0.100} for GRBL. Increase only if the magnet or pieces need more settle time. \\
\addlinespace
\texttt{SQUARE\_SIZE} & 1.5 in & Physical size of each board square. \\
\addlinespace
\texttt{BOARD\_SIZE} & 12.0 in & Physical size of the 8$\times$8 board area. \\
\addlinespace
\texttt{A1\_CENTER\_X/Y} & 5.000 / 4.420 in & Calibrated center of a1 in the G54 work frame. \\
\addlinespace
\texttt{CAPTURE\_VERTICAL\_SPACING} & 1.5 in & Storage slot row spacing; matches the board square size. \\
\bottomrule
\end{tabular}
\caption{Tunable hardware parameters}
\end{table}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
